---
title: "R Laboratory 2"
output: html_notebook
---

```{r}
library(ggplot2)
library(comprehenr)
```

**Exercise 1**

A set of measurements have been performed on the concentration of a contaminant
in tap water. The following tables reports a set of values (x), with the 
corresponding probabilities given by the two methods (p1 and p2)
Evaluate the expected values, E[X], and the variance, Var(X), for both methods



```{r}

x <- c(15.58, 15.9, 16, 16.1, 16.2)
p1 <- c(0.15, 0.21, 0.35, 0.15, 0.14)
p2 <- c(0.14, 0.05, 0.64, 0.08, 0.09)

cat("Expected value with method p1 ", sum(x*p1))
cat("Expected value with method p2 ", sum(x*p2))

cat("Variance with method p1 ", ( sum(p1*(x^2)) - (sum(x*p1))^2 ))
cat("Variance with method p2 ", ( sum(p2*(x^2)) - (sum(x*p2))^2 ))


```

**Exercise 2**

• the waiting time, in minutes, at the doctor’s is about 30 minutes, and the 
distribution follows an exponential pdf with rate 1/30

A) simulate the waiting time for 50 people at the doctor’s office and plot 
the relative
histogram

B) what is the probability that a person will wait for less than 10 minutes ?

C) evaluate the average waiting time from the simulated data and compare it with
the expected value (calculated from theory and by manipulating the probability
distributions using R)

D) what is the probability for waiting more than one hour before being received?



```{r fig.align="center", fig.width=8 , fig.height=6}
# exponential pdf
#Dt=1 min

#A)
waiting_time <- rexp(50, rate=1/30)

hist(waiting_time, col="steelblue", ylab="Occurences")
#-----------------
#B)
cat("The probability that a person will wait for less than 10 minutes is ", pexp(10 , rate=1/30)*100, "%")

#-----------------
#C)

exp_val_analysis <- mean(waiting_time)
expect_val_theo = 1/(1/30)   #from theory E[x] = 1/lambda
expect_val_distrib = integrate( function(x){ x*dexp(x, rate=1/30)}, 0 , +Inf)

cat("The expected value analysing the data is ", exp_val_analysis, ", from the theory ",expect_val_theo, ", from the distribution", expect_val_distrib[["value"]] ," ± ", round(expect_val_distrib[["abs.error"]], digits = 6))
#---------------
#D)
cat("The probability that a person will wait for more than 1 hour is ", (1 - pexp(60 , rate=1/30))*100, "%")
```

**Exercise 3**

• Let’s suppose that on a book, on average, there is one typo error every 
three pages.
If the number of errors follows a Poisson distribution, plot the pdf and cdf, 
and calculate the probability that there is at least one error on a 
specific page of the book.


```{r fig.align="center", fig.width=8 , fig.height=6}
#Dx=1 page

library(ggplot2)

rate <- 1/3
distr <- dpois(0:10, lambda=rate)
df <- data.frame((0:10) , distr) 
colnames(df) <- c('x' , 'y')

plot <- ggplot(df , aes(x=factor(x), y=y))
plot <- plot + geom_bar(stat='identity', fill="steelblue", color='grey3') + theme_linedraw()
plot <- plot + ggtitle('Poisson Distrib of Errors in the book') + labs(x="Number of Errors", y="Occurences") +                       theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12)) + scale_x_discrete(breaks = 0:10)
plot

#cumulative
cum_distr <- ppois(0:5, lambda=rate)
df2 <- data.frame((0:5) , cum_distr) 
colnames(df2) <- c('x' , 'y')

plot <- ggplot(df2 , aes(x=factor(x), y=y))
plot <- plot + geom_bar(stat='identity', fill="firebrick3", color='grey3', alpha=0.95) + theme_linedraw()
plot <- plot + ggtitle('Poisson Cumulative Distrib of Errors in the book') + labs(x="Number of Errors", y="Occurences") + theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12)) + scale_x_discrete(breaks = 0:5)
plot

#probability that there is at least one error on a specific page of the book
cat("The probability that there is at least one error on a specific page of the book is ", (1-distr[1])*100 , "%" )

```

**Exercise 4**

• we randomly draw cards from a deck of 52 cards, with replacement, until 
one ace is drawn. Calculate the probability that at least 10 draws are needed.


```{r}
ace_prob <- dbinom(x=0 , size=9, prob=4/52)

cat('The probability that are needed at least 10 draws to draw an ace is: ', ace_prob*100)
```

**Exercise 5**

• the time it takes a student to complete a TOLC-I University orientation and 
evaluation test follows a density function of the form

```{r}
         c(t − 1)(2 − t)    if  1 < t < 2
f (X) = 
          0             otherwise

```
          
where t is the time in hours.

a) using the integrate() R function, determine the constant c 
(and verify it analytically)

b) write the set of four R functions and plot the pdf and cdf, respectively

c) evaluate the probability that the student will finish the aptitude test
in more than 75 minutes. And that it will take 90 and 120 minutes.


```{r fig.align="center", fig.width=8 , fig.height=6}
funct <- function(t){(t-1)*(2-t)}
integral <- integrate(funct, lower = 1 , upper = 2)

cat('The constant c that correctly normalizes the integral is: ', 1/(integral$value), '\n')
cat('According with the analytical form of the integral is F(t) = -(t^3)/3 + 3/2*t^2 -2t, \nF(2)- F(1)= 6 is the value of the c constant')

#--------
#b)
dfunc <- function(t){
  distrib <- ifelse(t > 1 & t < 2, 6*(t-1)*(2-t),
                        0)
  return(distrib)
}

cfunc <- function(t){
  cdistr <- ifelse(t > 1 & t < 2, -2*(t^3) + 9*(t^2) -12*(t) + 5, ifelse(t>=2, 1, 0))
  return(cdistr)
}

x_val <- seq(1, 2 , length=20)



qfunc <- function(q){ 
  return(ifelse(q >=0 & q <= 1, uniroot(function(x){ return (c_func(x)-q)}, lower=1, upper=2)$root,
                "invalid input, t must be between 0 and 1"))
    }


rfunc <- function(n){
 return(to_vec(for(sampl in runif(n)) qfunc(sampl)))
  }



plot <- ggplot()
plot <- plot + geom_bar(aes(x=x_val, y=dfunc(x_val)), stat='identity', fill="steelblue", color='grey3') + theme_linedraw()
plot <- plot + ggtitle('Probability Distribution') + labs(x="x", y="y") +                                                      theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
plot



plot2 <- ggplot()
plot2 <- plot2 + geom_bar(aes(x=x_val, y=cfunc(x_val)), stat='identity', fill="firebrick2", color='grey3') + theme_linedraw()
plot2 <- plot2 + ggtitle('Cumulative Probability Distribution') + labs(x="x", y="y") +                       theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
plot2


prob1 <- integrate(dfunc, lower = 75/60 , upper =2)

cat('The probability that the test takes more then 75 minutes is: ', prob1$value*100, '% \n')
cat('The probability that the test takes exactly 90 minutes is: ', 0, '\nsingle point in a continous distribution\n')
cat('The probability that the test takes exactly 120 minutes is: ', 0, 
    '\n a single point in a continous distribution\n')

p = 1 - cfunc(1.5)
cat("The probability that the test takes between minute 90 and 120 is: ", p*100, "%")


```

**Exercise 6**

• the lifetime of tires sold by an used tires shop is 104 · x km, where x is a random
variable following the distribution function


```{r}
         2/x^2     if  1 < x < 2
f (X) =
          0       otherwise
```



          
a) write the set of four R functions and plot the pdf and cdf, respectively

b) determine the probability that tires will last less than 15000 km

c) sample 3000 random variables from the distribution and determine 
the mean value and the variance, using the expression Var(X) = E[X^2] − E[X]^2


```{r fig.align="center", fig.width=8 , fig.height=6}
funct2 <- function(x){ ifelse(x > 1 & x < 2, 2*(x^2),
                        0)}
integral2 <- integrate(funct2, lower = 1 , upper = 2)

cat('The integral value is ', (integral2$value), ' with an error of', integral2$abs.error)

dfunc2 <- function(x){
  distrib <- ifelse(x > 1 & x < 2,
                     2/(x^2),
                        0)
  return(distrib)
}

cfunc2 <- function(x){
  cdistr <- ifelse(x > 1 & x < 2, 
                   -2/x + 2, ifelse(x>=2, 1, 0))
  return(cdistr)
}


qfunc2 <- function(p){
    q_2 <- ifelse(p >= 0 & p <= 1,
               uniroot(function(x){ return (cfunc2(x) - p)}, lower=1, upper=2)$root,
               "invalid input, t must be between 0 and 1")
    return(q_2)
}


rfunc2 <- function(n){
  return(to_vec(for(sampl in runif(n)) qfunc2(sampl)))
}



x_val2 <- seq(1.01, 1.99 , length=20)



plot <- ggplot()
plot <- plot + geom_bar(aes(x=x_val2, y=dfunc2(x_val2)), stat='identity', fill="steelblue", color='grey3') + theme_linedraw()
plot <- plot + ggtitle('Probability Distribution') + labs(x="x", y="y") +                                                      theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
plot



plot2 <- ggplot()
plot2 <- plot2 + geom_bar(aes(x=x_val2, y=cfunc2(x_val2)), stat='identity', fill="firebrick2", color='grey3') + theme_linedraw()
plot2 <- plot2 + ggtitle('Cumulative Probability Distribution') + labs(x="x", y="y") +                       theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
plot2

#-----------
#b)
cat('The probability that tires will last less than 15000 km is ' , cfunc2(1.5)*100 , '%')


#-----------
#c)
n_samples <- 3000
samples <- rfunc2(n_samples)
tyre_life <- samples*(10^4)

avg <- sum(tyre_life)/n_samples
cat('The Average tire life of the tires sold by of this shop is', avg , 'km')

expec_2 <- sum(tyre_life^2)/n_samples

cat('The tire life Variance of the tires sold by this shop is', (expec_2 - (avg^2)) , 'km')

cat('Can estimate the lifetime of the tires sold by this shop as', avg ,'km', '±', sqrt(expec_2 - (avg^2)),'km')





```

