---
title: "R Laboratory 1"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---
---------------------------------------------------------------------------------

**Exercise 1 - Lakes of Scotland**

• The table that can be found at 
   https://en.wikipedia.org/wiki/List_of_lochs_of_Scotland
 gives the volume, area, length and maximum and mean depths of some Scottish
lakes. Create vectors, holding the lake’s name and all the parameters and build 
a dataframe called scottish.lakes from the vectors

1 evaluate the highest and lowest volume and area lake

2 order the frame with respect to the area and determine the two largest 
area lakes

3 by summing up the areas occpupied by the lakes, determine the area of Scotland 
covered by water

```{r}
scottish.lakes <- read.table('Data/LochLakes.csv', header=T , sep=',')

head(scottish.lakes)
#1 - evaluate the highest and lowest volume and area lake

indmax1 <- which.max(scottish.lakes[,'Volume.km..'])
indmax2 <- which.max(scottish.lakes[,'Area.km2.'])

indmin1 <- which.min(scottish.lakes[,'Volume.km..'])
indmin2 <- which.min(scottish.lakes[,'Area.km2.'])

cat("The lake with the highest Volume is ", paste(scottish.lakes[indmax1,1]), " with ", max(scottish.lakes[,'Volume.km..'])," km3")


cat("The lake with the highest Area is ", paste(scottish.lakes[indmax2,1]), " with ", max(scottish.lakes[,'Area.km2.'])," km2")


cat("The lake with the Lowest Volume is ", paste(scottish.lakes[indmin1,1]), " with ", min(scottish.lakes[,'Volume.km..'])," km3")


cat("The lake with the Lowest Area is ", paste(scottish.lakes[indmin2,1]), " with ", min(scottish.lakes[,'Area.km2.'])," km2")


#------------------------------------------------------------------------------------------

#2 - order the frame with respect to the area and determine the two largest area lakes

scottish_order <- scottish.lakes[order(scottish.lakes[,'Area.km2.'] , decreasing = TRUE) , ]
head(scottish_order)

cat("The two lakes with the largest Area are ", scottish_order[1,'Loch'], " and " , scottish_order[2,'Loch'], " respectively with an Area of ", scottish_order[1,'Area.km2.'] , " and " , scottish_order[2,'Area.km2.'], " km2")

#------------------------------------------------------------------------------------------

#3 - by summing up the areas occupied by the lakes, determine the area of Scotland covered by water

total_area <- sum(scottish.lakes[,'Area.km2.'])
cat('The total area of Scotland occupied by water is equal to ', total_area , 'km2')

```

------------------------------------------------------------------------------------

**Exercise 2 - Crude Oil Production**

• The following CSV file
https://drive.google.com/file/d/13WkinYhJNSr_XCldb3wpZOI0fgjer7Iw/view contains 
data on crude oil prices from 1861 to 2020, measured in US dollars per barrel.

1 Write R code that is able to read the file and import it in a data frame 
structure. (Hint: before loading the file, open it with a text editor 
and check its structure).

2 produce a plot with the Oil price as a function of the year

3 which is the highest price in history ? When did it occur ?

3 plot the derivative of the curve, simply evaluated with the finite difference
formula (forward derivative):    dprice / dyear = pricej+1 − pricej



```{r}
#1 - read data
oil <- read.csv('Data/crude-oil-prices.csv')
colnames(oil) <- c("Entity", "Code", "Year", "Price")   #more accessible

#2 - produce a plot with the Oil price as a function of the year
library(ggplot2)

plot <- ggplot(oil, aes(x=Year , y=Price)) + theme_linedraw()
plot <- plot + ggtitle('Crude Oil Price in $/barrel over Years')
plot <- plot + labs(x="Year", y="Price($/barrel)")
plot <- plot + theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12)) 
plot + geom_line(color="red")
#--------------------------
#3 - which is the highest price in history? when did it occur?

ind_pricemax <- which.max(oil[,'Price'])

cat("The year with the highest price of crude oil is ", paste(oil[ind_pricemax,'Year']), " with ", max(oil[,'Price']),"$/barrel")
#--------------------------
#4 - plot the derivative of the curve

price_deriv <- c(diff(oil[,'Price']), NA)

plot2 <- ggplot(oil[1:nrow(oil)-1,], aes(x=Year , y=price_deriv[-length(price_deriv)]))
plot2 <- plot2 + theme_linedraw() + ggtitle('Oil price derivative over time')
plot2 <- plot2 + labs(x="Year", y="Price derivative")
plot2 <- plot2 + theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12)) 
plot2 + geom_line(color="navy")

```


---------------------------------------------------------------------------------

**Exercise 3 - World Coal Production**

• The following CSV file
https://drive.google.com/file/d/1U1hK5o_d3vl8twwGaRJNcDJ79O9I5zJ1/view?usp=sharing

contains data on the coal production (in TW · hour) for several countries 
in the World.

1 Write R code that is able to read the file and import it in a tibble structure

2 count the number of countries available in the file and produce a barplot
with the number of entries for each country

For the following items select only the years ≥ 1970:
3 selecting only the year after 1970, determine the total integrated production 
for each country and print the top 5 Countries with highest coal productions

4 for the 5 top Countries, create a plot of production as a function of time

5 generate a plot with the cumulative sum of the World’s coal production
over the years


```{r fig.align="center", fig.width=8 , fig.height=5}
#1 - read coal-production-by-country.csv and import in a tibble structure
library(tibble)

coal = tibble(read.csv("Data/coal-production-by-country.csv"))
colnames(coal) <- c("Entity", "Code", "Year", "CoalProdTWh")   #more accessible
head(coal)
#----------------------
#2 - count the number of countries available in the file and produce a barplot

countries <- unique(coal[,'Entity'])
cat("Counting , the number of avalaible countries in the file are " , nrow(countries))

library(ggplot2)

plot <- ggplot(coal, aes(x=Entity)) + theme_linedraw()
plot <- plot + ggtitle('Number of entries for each country')
plot <- plot + labs(x="Countries", y="Occurences") 
plot <- plot + theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10, angle = 90, hjust = 1, vjust = 0.5), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12),
                  aspect.ratio = 0.2)
plot <- plot + scale_x_discrete(breaks =unique(coal$Entity)[c(TRUE, rep(FALSE, 5))])
plot + geom_bar(width = 0.75, fill="steelblue")
#----------------------
#3 - only the years ≥ 1970: determine the total integrated production for each 
#    country and print the top 5 Countries with highest coal productions

coal_ <- coal[coal[,'Year'] >= 1970,]

sum_by_country = aggregate(coal_[,'CoalProdTWh'], coal_[,'Entity'], sum)

sum_df <- data.frame(sum_by_country[1], sum_by_country[2])                   
colnames(sum_df) <- c("Country", "Tot_Prod")

sum_df = sum_df[order(sum_df$Tot_Prod, decreasing=TRUE), ]
row.names(sum_df) <- NULL

print(sum_df[1:5,])

not_countries <- c("World", "Asia Pacific", "Asia and Oceania", "OECD", "North America", "Eurasia", "Europe", "EU-28", "CIS", "South & Central AMerica", "USSR")

top_5 = sum_df[!sum_df$Country %in% not_countries, ]$Country[1:5]
sum_df = sum_df[!sum_df$Country %in% not_countries, ][1:5, ]
row.names(sum_df) <- NULL
cat("The top 5 countries for TWh Coal production")
print(sum_df)

#---------------------------
#4 - for the 5 top Countries, create a plot of production as a function of time

coal_top_5 = coal_[coal_$Entity %in% top_5, ]


plot <- ggplot(coal_top_5, aes(x=Year , y=CoalProdTWh)) + theme_linedraw()
plot <- plot + ggtitle('Coal Production for Top5 countries over Years after 1970')
plot <- plot + labs(x="Year", y="Coal Production [TWh]")
plot <- plot +  geom_line(aes(color=Entity), lwd=1.1)
plot <- plot + theme(plot.title = element_text(size=12), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12)) 
plot
#---------------------------
#5 - generate a plot with the cumulative sum of the World’s coal
#     production over the years

library(dplyr)

world_prod <- filter(coal_, coal_$Entity == 'World')
world_cumsum <- unlist(cumsum(world_prod[,"CoalProdTWh"]))
world_prod['CumulatSum'] <- world_cumsum



plot <- ggplot(world_prod, aes(x = Year, y=world_cumsum)) + theme_linedraw()
plot <- plot + ggtitle('World cumulative coal production over time')
plot <- plot + labs(x="Year", y="Coal Production [TWh]")
plot <- plot + geom_line(lwd=1.5, color="darkgreen")
plot <- plot + theme(plot.title = element_text(size=12), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12)) 
plot     

```

---------------------------------------------------------------------------------

**Exercise 4 - Covid19 Vaccine Data**


• Data on Covid-19 vaccine can be found on.
• The file vaccinationsby-manufacturer.csv,
(https://github.com/owid/covid-19-data) contains a summary
on the vaccination data grouped by Vaccine type and Country.

1 filter() the original tibble by selecting the following countries: Italy

2 plot the number of vaccines given as a function of time for the 
different vaccine manufacturer

3 from the same tibble plot the total number of vaccines shot per day in Italy

4 do the same exercise for the following countries: Germany and United States of America


```{r fig.align="center", fig.width=8 , fig.height=5}
library(tibble)
library(dplyr)
 
cov_data <- tibble(read.csv('Data/covid.csv'))
cov_data$date <- as.Date(cov_data$date)  # convert date column as date type

#1 - filter() the original tibble by selecting the following countries: Italy
cov_ita <- filter(cov_data, cov_data$location == 'Italy')
print(head(cov_ita))
#------------------
#2 - plot the number of vaccines given as a function of time for the different vaccine manufacturer

library(ggplot2)

plot <- ggplot(cov_ita, aes(x=date , y=total_vaccinations , color = vaccine))
plot <- plot + theme_linedraw() + ggtitle('Number of Vaccines given vs Time - Italy')
plot <- plot + labs(x="Date", y="Number of Vaccines given")
plot <- plot + theme(plot.title = element_text(size=16), 
            axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
            axis.text.y = element_text(size=10),
            axis.title.x = element_text(size=12),
            axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")                                                               + scale_y_continuous(trans='log10') + 
                                                                                geom_line(lwd = 1.1)
plot

#------------------
#3 - from the same tibble plot the total number of vaccines shot per day in Italy

vaccines_per_day <- rep(0, length(unique(cov_ita$date)))
for (i in 1:length(unique(cov_ita$date))){
    date_ = unique(cov_ita$date)[[i]]
    for (vaccine_ in unique(cov_ita$vaccine)){
        current_shots  = filter(cov_ita, date==date_ & vaccine==vaccine_)
        previous_shots = filter(cov_ita, date==(date_-1) & vaccine==vaccine_)
        if (dim(current_shots)[1]!=0){
            if (dim(previous_shots)[1]!=0){
                vaccines_per_day[i] <- vaccines_per_day[i] + current_shots$total_vaccinations - previous_shots$total_vaccinations
            }
        }
    }
}


plot2 <- ggplot()
plot2 <- plot2 + theme_linedraw() + ggtitle('Total Number of Vaccines given per day vs Time - Italy')
plot2 <- plot2 + labs(x="Date", y="Number of Vaccines given")
plot2 <- plot2 + theme(plot.title = element_text(size=16), 
              axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
              axis.text.y = element_text(size=10),
              axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")+                   geom_line(aes(x=unique(cov_ita$date) , y=vaccines_per_day), lwd = 1.1 , color='red')
plot2





```
```{r  fig.align="center", fig.width=8 , fig.height=5}

#------------------
#4 - do the same exercise for Germany 

library(tibble)
library(dplyr)
library(ggplot2)

cov_ger <- filter(cov_data, cov_data$location == 'Germany')

plot <- ggplot(cov_ger, aes(x=date , y=total_vaccinations , color = vaccine))
plot <- plot + theme_linedraw() + ggtitle('Number of Vaccines given vs Time - Germany')
plot <- plot + labs(x="Date", y="Number of Vaccines given")
plot <- plot + theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y") + scale_y_continuous(trans='log10') + geom_line(lwd = 1.1)
plot

#no problems in data
total_vaccines_per_day <- aggregate(cov_ger$total_vaccinations, list(cov_ger$date),sum)[[2]]
vaccines_per_day <- diff(total_vaccines_per_day)      

plot2 <- ggplot()
plot2 <- plot2 + theme_linedraw() + ggtitle('Total Number of Vaccines given per day vs Time - Germany')
plot2 <- plot2 + labs(x="Date", y="Number of Vaccines given")
plot2 <- plot2 + theme(plot.title = element_text(size=16), 
              axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
              axis.text.y = element_text(size=10),
              axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")+                   geom_line( aes(x=unique(cov_ger$date)[-1] , y=vaccines_per_day), lwd = 1.1 , color='steelblue')
plot2

```


```{r fig.align="center", fig.width=8 , fig.height=5}
#------------------
#4 - do the same exercise for the United States of America

library(tibble)
library(dplyr)
library(ggplot2)

cov_usa <- filter(cov_data, cov_data$location == 'United States')

plot <- ggplot(cov_usa, aes(x=date , y=total_vaccinations , color = vaccine))
plot <- plot + theme_linedraw() + ggtitle('Number of Vaccines given vs Time - United States')
plot <- plot + labs(x="Date", y="Number of Vaccines given")
plot <- plot + theme(plot.title = element_text(size=16), 
              axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
              axis.text.y = element_text(size=10),
              axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")                      + scale_y_continuous(trans='log10') + geom_line(lwd = 1.1)

plot

#no problems in data??
total_vaccines_per_day <- aggregate(cov_usa$total_vaccinations, list(cov_usa$date),sum)[[2]]
vaccines_per_day <- diff(total_vaccines_per_day)      

plot2 <- ggplot()
plot2 <- plot2 + theme_linedraw() + ggtitle('Total Number of Vaccines given per day vs Time - United States')
plot2 <- plot2 + labs(x="Date", y="Number of Vaccines given")
plot2 <- plot2 + theme(plot.title = element_text(size=16), 
            axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
            axis.text.y = element_text(size=10),
            axis.title.x = element_text(size=12),
            axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")+                   geom_line( aes(x=unique(cov_usa$date)[-1] , y=vaccines_per_day), lwd = 1.1 , color='steelblue')
plot2


```
Observing these plots there is a problem in the data: around March 2022
the registered number of vaccines given of Johnson&Johnson leads to a 
negative difference in the Total number of vaccines given... 
Try to adjust this...


```{r fig.align="center", fig.width=8 , fig.height=5}

library(tibble)
library(dplyr)
library(ggplot2)

df <- cov_usa[cov_usa$date > as.Date("2022-03-01"), ]
for (i in 1:length(df$date)){
    if(df[[i, "total_vaccinations"]] < mean(df[df$vaccine == "Johnson&Johnson", ]$total_vaccinations)){
        date_ = df[[i, "date"]]
        }
}

# fix the error replacing the bad record with it previous value

bad_record <- which(cov_usa$date == date_ & cov_usa$vaccine == "Johnson&Johnson")
cov_usa[bad_record, ] <- cov_usa[cov_usa$date == date_-1 & cov_usa$vaccine == "Johnson&Johnson", ]


# filter to repair eventual differences
difference_filter <- function(difference){
    for (i in 1:length(difference)){
        if (difference[i]<0){
            difference[i]=0
        }
        if (difference[i]>10*mean(difference)){
            difference[i]=0
        }
    }
    return(difference)
}


plot <- ggplot(cov_usa, aes(x=date , y=total_vaccinations , color = vaccine))
plot <- plot + theme_linedraw() + ggtitle('Number of Vaccines given vs Time - United States')
plot <- plot + labs(x="Date", y="Number of Vaccines given")
plot <- plot + theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")                      + scale_y_continuous(trans='log10') + geom_line(lwd = 1.1)

plot


total_vaccines_per_day <- aggregate(cov_usa$total_vaccinations, list(cov_usa$date),sum)[[2]]
vaccines_per_day <- difference_filter(diff(total_vaccines_per_day))      

plot2 <- ggplot()
plot2 <- plot2 + theme_linedraw() + ggtitle('Total Number of Vaccines given per day vs Time - United States')
plot2 <- plot2 + labs(x="Date", y="Number of Vaccines given")
plot2 <- plot2 + theme(plot.title = element_text(size=16), 
              axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
              axis.text.y = element_text(size=10),
              axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")+                   geom_line( aes(x=unique(cov_usa$date)[-1] , y=vaccines_per_day), lwd = 1.1 , color='firebrick')
plot2

```

• The file vaccinations.csv contains country-by-country data on global COVID-19
vaccinations. The dataset includes some subnational locations 
(England, Northern Ireland, Scotland, Wales, . . .) and international aggregates
(World, continents, European Union ldots). 

They can be identified by their iso code that starts with OWID

1 selecting all the European countries in the tibble, plot the number 
of daily vaccinations per million as a function of date


```{r fig.align="center", fig.width=8 , fig.height=5}
library(tibble)
library(dplyr)
library(ggplot2)

vaccinations_data <- tibble(read.csv('Data/vaccinations.csv'))
vaccinations_data$date <- as.Date(vaccinations_data$date)  # convert date column as date type

tail(vaccinations_data)

#1 selecting all the European countries in the tibble, plot the number 
#of daily vaccinations per million as a function of date
european_vac_df <- vaccinations_data[vaccinations_data$iso_code == "OWID_EUR" ,]

total_vaccines_per_day <- aggregate(european_vac_df$daily_vaccinations_per_million, list(european_vac_df$date), sum)

plot <- ggplot()
plot <- plot + theme_linedraw() + ggtitle('Daily Vaccination given per milion vs Time - Europe')
plot <- plot + labs(x="Date", y="Number of Daily Vaccines per milion")
plot <- plot + theme(plot.title = element_text(size=16), 
              axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
              axis.text.y = element_text(size=10),
              axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")+                   geom_line( aes(x=european_vac_df$date , y=european_vac_df$daily_vaccinations_per_million), lwd = 1.1 , color='darkorange2')
plot


```

2 study the data structure and produce few relevant plots of your taste

```{r fig.align="center", fig.width=8 , fig.height=6}
colnames(vaccinations_data) #exploring

entities <- c("Europe", "Asia", "Africa", "North America", "South America", "Australia", "Italy")
vacc_df_locations <- vaccinations_data[vaccinations_data$location %in% entities, ]

library(ggplot2)


plot <- ggplot(data=vacc_df_locations)
plot <- plot + theme_linedraw() + ggtitle('People Vaccinated per Hundred vs Time - Different Locations')
plot <- plot + labs(x="Date", y="People Vaccinated per Hundred")
plot <- plot + theme(plot.title = element_text(size=16), 
              axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
              axis.text.y = element_text(size=10),
              axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")+                   geom_line( aes(x=date , y=people_vaccinated_per_hundred , color =location), lwd = 1.1)
plot



plot <- ggplot(data=vacc_df_locations)
plot <- plot + theme_linedraw() + ggtitle('People Fully Vaccinated per Hundred vs Time - Different Locations')
plot <- plot + labs(x="Date", y="People Fully Vaccinated per Hundred")
plot <- plot + theme(plot.title = element_text(size=16), 
              axis.text.x = element_text(size=10, angle = 60, hjust = 1), 
              axis.text.y = element_text(size=10),
              axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) + scale_x_date(date_breaks = "1 months",date_labels="%b-%y")+                   geom_line( aes(x=date , y=people_fully_vaccinated_per_hundred , color =location), lwd = 1.1)
plot



```
From the official global database WHO-COVID-19-global-data.csv, try to analyse the effectiveness of vaccines in Italy


```{r fig.align="center", fig.width=15 , fig.height=6}
library(tibble)
library(dplyr)
library(ggplot2)

glob_covid_df = tibble(read.csv("Data/WHO-COVID-19-global-data.csv"))

# convert date column as date type
glob_covid_df$Date_reported = as.Date(glob_covid_df$Date_reported)

tail(glob_covid_df)

# selecting Italy
covid_df.italy = glob_covid_df[glob_covid_df$Country =="Italy", ]
vaccinations_df.italy = vaccinations_data[vaccinations_data$location == "Italy", ]


upper_y = 250000
cols    = c("new_cases"="firebrick2", "people_vaccinated_per_hundred"="steelblue4")
plot <- ggplot() +   
     theme_linedraw() +  
     # new cases line                                          
     geom_line(data=covid_df.italy,
               aes(x=Date_reported, y=New_cases, colour="new_cases"),
               lwd=1.1) +
     # people vaccinated line 
     geom_line(data=vaccinations_df.italy,
               aes(x=date, y=people_vaccinated_per_hundred*upper_y/100, colour="people_vaccinated_per_hundred"),
               lwd=1.2) +         

     xlab("date") +
     ggtitle("New cases & People vaccinated vs time - Italy") +
     theme(plot.title = element_text(size=14 , hjust = 0.5),
           axis.text.x = element_text(size=10, angle = 60, hjust = 1)) +
     scale_x_date(date_breaks = "1 months", date_labels="%b-%y") +
     scale_color_manual(name="Legend",
                        values=cols) +
    
     theme(axis.text.y = element_text(size=10),
             axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) 
     
    # adjusting y axis
     scale_y_continuous(
           name = "New cases",
           limits = c(0, upper_y),
          sec.axis = sec_axis(~.*100/upper_y, name="People vaccinated per hunderd")) 

plot
```
The new cases are completely uncorrelated to the vaccinated people.   
Even if the vaccinated people increased, indeed, the new cases do not decrease.
The vaccines do not prevent infections and transmissions...

```{r fig.align="center", fig.width=15 , fig.height=6}
upper_y = 1000


cols    = c("new_deaths"="firebrick2", "people_vaccinated_per_hundred"="steelblue4")

p1 <- ggplot() +   
     theme_linedraw() +  
     # new cases line                                          
     geom_line(data=covid_df.italy,
               aes(x=Date_reported, y=New_deaths, colour="new_deaths"),
               lwd=1.5) +
     # people vaccinated line 
     geom_line(data=vaccinations_df.italy,
               aes(x=date, y=people_vaccinated_per_hundred*upper_y/100, colour="people_vaccinated_per_hundred"),
               lwd=1.5) +         
     xlab("date") +
     ggtitle("New deaths & People vaccinated vs time - Italy") +
     theme(plot.title = element_text(size=14 , hjust = 0.5),
           axis.text.x = element_text(size=10 , angle = 60, hjust = 1)) +
     scale_x_date(date_breaks = "1 months", date_labels="%b-%y") +
     scale_color_manual(name="Legend",
                        values=cols) +
    
        
     theme(axis.text.y = element_text(size=10),
             axis.title.x = element_text(size=12),
              axis.title.y = element_text(size=12)) +
      
    
     # adjusting y axis
     scale_y_continuous(
           name = "New deaths",
           limits = c(0, upper_y),
           sec.axis = sec_axis(~.*100/upper_y, name="People vaccinated per hunderd")) 

p1

```
Otherwise the number of vaccinated people and the deaths by Covid-19 are 
strictly related, from this plot in fact one can see that when people start to
vaccinated largely, the number of new deaths enormously decreases.

