---
title: "R Laboratory 3"
output: html_notebook
---

```{r}
library(ggplot2)
```



## Exercise 1

• The triangular distribution, in the interval (a, b), is given by the following:

$$    
f (X) =     
  \begin{cases}
  \frac{2(x − a)}{(b − a)(c − a)} & if \quad a \leq x < c \\
  \frac{2(b − x)}{(b − a)(b − c)} & if \quad c \leq x \leq b \\
   0 & \text{otherwise}
\end{cases}
$$

a) plot the function, given the interval (a, b)

b) and write an algorithm to generate random numbers from the triangular distr.

c) generate 10^4 random number from the distribution, show them in an histogram
and superimpose the analytical curve


```{r fig.align="center", fig.width=8 , fig.height=6}

triang <- function(x,a=-3,b=3,c=0){
  ifelse((x < c & x >= a) , 2*(x-a)/((b-a)*(c-a)),
         ifelse((x <= b & x >= c) , 2*(b-x)/((b-a)*(b-c)), 0))
}

a <- -4
b <- 4

x_ax <- seq(a,b , length=103)
y_ax <- triang(x_ax , a , b)

plot <- ggplot() + theme_linedraw() + 
                  geom_line(aes(x=x_ax, y=y_ax), stat='identity', color="steelblue", lwd=2)
plot <- plot + ggtitle('Triangular Distribution') + labs(x="x", y="y") +                                                            theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
plot

#----------
#b)

cum_tr <- function(x,a,b,c){
    integral <- numeric(length(x))
    for (i in seq_along(x)){
      integral[i] <- integrate(triang, lower=a , upper = x[i], a,b)$value
    }
    return(integral)
   }

x_ax2 <- seq(a,b, length=100)
y_ax2 <- cum_tr(x_ax2, a, b, 0)
 
plot <- ggplot() + theme_linedraw() + 
                  geom_line(aes(x=x_ax2, y=y_ax2), stat='identity', color="firebrick3", lwd=2)
plot <- plot + ggtitle('Cumulative Distribution') + labs(x="x", y="y") +                                                            theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
plot

inverse <- function(y, a,b,c){
  uniroot(function(x){ cum_tr(x,a,b,c) - y }, lower=-4 , upper=4)$root
}

# Quantile function
q_triang <- function(y, a,b,c, lower.value=-4, upper.value=4){
  output <- numeric(length(y))
  
  for (i in seq_along(y)){
    if (y[i]==1) output[i] <- upper.value      
    else if (y[i]==0) output[i] <- lower.value
    else output[i] <- inverse(y[i],a,b,c)
  }
  return(output)
}


rand_tr <- function(n, a, b, c){
  return(q_triang(runif(n), a,b,c))
}


pdf <- rand_tr(10^4, a, b, 0)

plot <- ggplot() + theme_linedraw() + 
                  geom_histogram(aes(x=pdf, y=..density..), fill='navy', alpha=0.7, bins = 50, color='black') +
                  geom_line(aes(x=x_ax, y=y_ax), stat='identity', color="firebrick3", lwd=2)
plot <- plot + ggtitle('Triangular pdf superimposed with analytical form') + labs(x="x", y="y") +                                   theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
plot
  
  
```

## Exercise 2 - Markov’s inequality

• Markov’s inequality represents an upper bound to probability distributions:

$$
             
P(X \geq k) \leq \frac{E[X]}{k} \quad    for \   \ k > 0
        
$$

               
• having defined a function         $$  G(k) = 1 − F(k) ≡ P(X \geq k) $$


plot G(k) and the Markov’s upper bound for

a) the exponential, Exp(λ = 1), distribution function

b) the uniform, U (3, 5), distribution function

c) the binomial, Bin(n = 1, p = 1/2), distribution function

d) a Poisson, Pois(λ = 1/2), distribution function


```{r fig.align="center", fig.width=10 , fig.height=7}

G <- function(cdf){
  return (1 - cdf)
}

markov_ub <- function(k, expec_val){
  return(expec_val/k)
}

pl_funct <- function(k_val , cdf , expec_val, pl_title="Markov's Inequality", xlab='x'){

   ggplot() + theme_linedraw() + 
       geom_line(aes(x=k_val, y=markov_ub(k,expec_val), color='Markov\'s UpperBound'), linetype="dashed", lwd=1.5) +
       geom_line(aes(x=k_val, y=G(cdf), color="G(k)"), lwd=1.5) +
     labs(color='') +  theme(legend.text=element_text(size=12)) +
        ggtitle(pl_title)+ labs( x=xlab, y="y") +  
             theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
}


k <- seq(1, 10, length = 100)
pl_funct(k, cdf=pexp(k, 1),expec_val=1, pl_title="Markov's Inequality for Exp Distribution with lambda=1")   + xlim(0,10)

k<- seq(0,7, length=100)
pl_funct(k, cdf=punif(k, 3,5),expec_val=(3+5)/2, pl_title="Markov's Inequality for Uniform Distribution (3,5)") + xlim(0,8) + ylim(0,2.5)

k<- seq(0,7, length=100)
pl_funct(k, cdf=pbinom(k , 1, 1/2),expec_val=1*1/2, pl_title="Markov's Inequality for Binomial Distribution (n=1 , prob=1/2)")  + xlim(0,7.3)  + ylim(0,2.5)

k<- seq(0,7, length=100)
pl_funct(k, cdf=ppois(k, 1/2),expec_val=1/2, pl_title="Markov's Inequality for Poisson Distribution with lambda=1/2") + xlim(0,7.3) + ylim(0,2.5)

```


## Exercise 3 - Chebyshev’s inequality

• Chebyshev’s inequality tell us that

$$
P( | X − μ | \geq kσ) \leq \frac{1}{k^2}
$$

• which can also be written as

$$
P( |X − μ |< kσ) \geq 1 − \frac{1}{k^2}
$$

• use R to show, with a plot, that Chebyshev’s inequality is is an upper bound to the
following distributions:

a) a normal distribution, N(μ = 3, σ = 5)

b) an exponential distribution, Exp(λ = 1)

c) a uniform distribution U (1 − √2, 1 + √2)

d) a Poisson, Pois(λ = 1/3), distribution function


```{r fig.align="center", fig.width=10 , fig.height=7}


cum_ksigma <- function(k , cdf, mu, sigma){
  return(cdf(mu - k*sigma) + 1 -cdf(mu + k*sigma))
}

cheb_ub <- function(k){
  return((1/k^2))
}




pl_funct2 <- function(x_val , cdf , mu , sigma, pl_title="Chebyshev's Inequality", xlab='x'){

   ggplot() + theme_linedraw() + 
       geom_line(aes(x=x_val, y=cheb_ub(x_val) , color=' Chebyshev\'s Bound'),  linetype="dashed" , lwd=1.5) +
       geom_line(aes(x=x_val, y=cum_ksigma(x_val,cdf,mu,sigma), color='P_kSigma'), lwd=1.5) +
        ggtitle(pl_title)+ labs( x=xlab, y="y") + labs(color='') +                                               theme(legend.text=element_text(size=12)) +
             theme(plot.title = element_text(size=16), 
                  axis.text.x = element_text(size=10), 
                  axis.text.y = element_text(size=10),
                  axis.title.x = element_text(size=12),
                  axis.title.y = element_text(size=12))
}

x_ax <- seq(0, 10, by=0.1)
mean <- 3
sigma <- 5
pl_funct2(x_ax, function(x) pnorm(x, mean=mean, sd=sigma), mu=mean, sigma=sigma, pl_title="Chebyshev's Inequality for Normal Distribution (3,5) ")  + ylim(0,4)


lambda <- 1
mean <- 1/lambda
sigma <- 1/lambda^2
pl_funct2(x_ax, function(x) pexp(x, rate=lambda) , mu=mean, sigma=sigma, pl_title="Chebyshev's Inequality for Exponential Distribution lambda=1 ")  + ylim(0,3)


minim <- 1 - sqrt(2)
maxim <- 1 + sqrt(2)
mean <- (maxim + minim ) / 2
sigma <- sqrt((maxim - minim)^2/12)
pl_funct2(x_ax, cdf= function(x) punif(x, min=minim, max=maxim) , mu=mean, sigma=sigma, pl_title="Chebyshev's Inequality for Uniform Distribution (1 − √2, 1 + √2)) ")  + ylim(-0.2,2) 

lambda <- 1/3
mean <- lambda
sigma <- sqrt(lambda)
pl_funct2(x_ax, cdf= function(x) ppois(x, lambda=lambda) , mu=mean, sigma=sigma, pl_title="Chebyshev's Inequality for Poisson Distribution (lambda = 1/3) ")  + ylim(-0.2,2) 

```